Efficient GPU implementation of multicomponent and multiphase Lattice Boltzmann models.

### 2010-04-04 paper.tex
Break lines in 3 more paragraphs.
Rewrite parts of the introduction section to make it shorter.
Also add info about Succi's GPU paper.

%%%% Start %%%%%%
\section{Introduction}
\label{sec:introduction}
During the last two decades since its introduction \cite{mcnamara}, the Lattice Boltzmann Method (LBM)
became mature enough to compete with other well-established approaches to computational fluid dynamics (CFD).
Due to its kinetic nature, LBM can be used to simulate not only simple
hydrodynamics but also thermal flows \cite{karlin-minimalmodels,yuan-thermal},
micro-flows \cite{ansumali-small-knudsen}, ferrofluids \cite{rosensweig,kuzmin-aniso}, and
multiphase flows \cite{rothman-color,Shan-chen:extended,swift}. In the presence of complex
physical effects such as turbulence, multi-phase flows, flow of oil and gas through porous
material, flow around surfaces with complex geometries, or blood flow with deformable
particles, the solution of the basic Navier-Stokes equations becomes difficult or
impractical. On the other hand, micro-scale calculations for such systems based
on the Molecular Dynamics approach are computationally extremely demanding.
The Lattice Boltzmann Method (LBM), which focuses on the meso-scale, may be the best
choice for such complex flow problems, being able to capture the salient details of both macro- and micro-flows.

{\color{red} Michal, please add here more details.}
The lattice Boltzmann method is the explicit numerical method opposing some restrictions on the time propagation step. Thus the physical time step corresponding to the numerical time step is relatively small. Performing large-time-scale simulations become unpractical due to the high computational demand. To overcome the problem of the high computational demand the physical problems are calculated with the help of parallel processors of a cluster. The recent years trend is to perform the simulations on Graphical Processing Units (GPUs) which can be viewed as small supercomputers due to the large number of independent threads performing a simulation. The attractive part of simulations on GPUs is the low price (in comparison with clusters) and large performance. Not all the problems are matched to be simulated on parallel machines. However, the lattice Boltzmann method is a perfect match for parallel structure machines as GPUs.

During the last years the method was applied for many different problems, such as, to name a few, the calculation of the drag coefficient for the two-dimensional flow through cylinders \cite{tolke-twod} and moving sphere in three-dimensional channel \cite{tolke-GPU}. To the authors' best knowledge only simple hydrodynamics problems GPUs implementation is thoroughly described in the literature. However, the nearest neighbours (NN) interaction which includes advanced boundary lattice Boltzmann methods and multiphase problems are still not covered. For many multiphase problems the local density gradient is required to properly simulate the physics. The implementation of the multiphase models are more challenging than simple hydrodynamics problems because of the additional information required from other nodes. The simple hydrodynamics lattice Boltzmann method is local and is perfectly parallelized. The NN problems requires the additional information from other nodes and can't be addressed as local any more.

This work's aim is to present the efficient GPUs algorithms to implement the binary liquids models based on the free-energy approach and the Shan-Chen method. First, we thoroughly describe the GPUs algorithms. Then the Shan-Chen model and the free-energy binary liquid model lattice Boltzmann implementation is presented. The numerical results for the comparison benchmark between those two models is done in terms of performance and accuracy in the ``Numerical results`` section. We conclude the paper with the main findings.

@@

\section{GPU implementation}
\section{Results}
\section{Conclusion}
%%%% Acknowledgments %%%%%%%%
\section*{Acknowledgments}
A.~Kuzmin wants to thank the Alberta Ingenuity Fund for their financial support.

>>>>

%%%% Start %%%%%%
\section{Introduction}
\label{sec:introduction}
During the last two decades since its introduction \cite{mcnamara}, the Lattice Boltzmann Method (LBM)
became mature enough to compete with other well-established approaches to computational fluid dynamics (CFD).
Due to its kinetic nature, LBM can be used to simulate not only simple
hydrodynamics but also thermal flows \cite{karlin-minimalmodels,yuan-thermal},
micro-flows \cite{ansumali-small-knudsen}, ferrofluids \cite{rosensweig,kuzmin-aniso}, and
multiphase flows \cite{rothman-color,Shan-chen:extended,swift}. In the presence of complex
physical effects such as turbulence, multi-phase flows, flow of oil and gas through porous
material, flow around surfaces with complex geometries, or blood flow with deformable
particles, the solution of the basic Navier-Stokes equations becomes difficult or
impractical. On the other hand, micro-scale calculations for such systems based
on the Molecular Dynamics approach are computationally extremely demanding.
The lattice-Boltzmann method, which focuses on the meso-scale, may be the best
choice for such complex flow problems, being able to capture the salient details of both macro- and micro-flows.

LB is an explicit numerical method, which poses some restrictions on the size of the time step --
it has to be kept relatively small in order to preserve the numerical stability of the scheme.  Long-time
simulations in large domains can thus quickly become impractical due to the high computational demand.
The usual solution to this problem is to use a cluster to run the simulations.  In the last few
years, an interesting alternative in the form of Graphics Processing Units of commodity video
cards has emerged.  The GPUs can be viewed as small supercomputers due to their high
computational power (e.g. 1 TFLOPS for a Tesla C1060) and ability to execute multiple
independent threads simultaneously.  Their low price makes them an interesting hardware
platform for numerical calculations of many kinds, but it has to be kept in mind that not
all problems are well suited for the specific architecture of the GPU.  The Lattice-Boltzmann
method, due it its inherent parallelism and locality is however a perfect match for this
new hardware environment.

This was quickly noticed by the CFD community and in the last few years the method was
implemented on NVIDIA GPUs and applied to many different problems, such as calculation of
the drag coefficient for the two-dimensional flow through cylinders \cite{tolke-twod} and
for a moving sphere in a three-dimensional channel \cite{tolke-GPU}.  In the majority of these
problems only fully local, single-fluid LB models were used.  Only very recently some
work on more complex models with NN interactions has been done {\color{red} cite Succi's paper},
but using a code ported from a CPU version, and not optimized specifically for the GPU
environment.
%%
% To the authors' best knowledge only simple
% hydrodynamics problems GPUs implementation is thoroughly described in the literature.
%%
% This is actually not true anymore -- see the Succi's PRE paper about LBM on GPUs
The aim of this paper is to present efficient GPU-optimized algorithms for LB models
with NN interaction terms.  Such models are abundant and of enormous practical importance.
They include the Shan-Chen model, free-energy binary fluid models and higher-order
boundary conditions for fluid flows.  In the following sections, we first present
the details of the Shan-Chen and free-energy LB models.  The reader
is then briefly introduced to the main features of the CUDA environment
that we use to program the GPUs, which is followed by a thorough description
of the GPU algorithms and an analysis of their efficiency.  In section \ref{sec:bechmark}, the two
binary fluid models are compared in terms of performance and accuracy.
We conclude the paper with a summary of the main findings.

@@

\section{GPU implementation}
\section{Results}
\label{sec:benchmark}

\section{Conclusion}
%%%% Acknowledgments %%%%%%%%
\section*{Acknowledgments}
A.~Kuzmin wants to thank the Alberta Ingenuity Fund for their financial support.
